Experimentally informed site-specific substitution models deepen phylogenetic estimates of the divergence of viral lineages

### 2018-02-26 manuscript/manuscript.tex
rewrite of phylobayes section

Finally, we compared the branch lengths estimated between two different models with site-specific stationary states. 
We compared the branch lengths estimated by ExpCM+$\Gamma\omega$ (avg) with the branch lengths estimated by the mutation-selection model implemented in \texttt{phylobayes}. 
The \texttt{phylobayes} mutation-selection model estimates the amino-acid preferences which define the stationary state in a Bayesian framework rather than defining them from the DMS \textit{a priori}. 
Branch length estimates are very similar between the ExpCM+$\Gamma\omega$ (avg) and the \texttt{phylobayes} mutation-selection model. 
The equivalence between the two models shows the general importance of site-specific stationary states. 
The local accuracy of the DMS preferences and the ExpCM can still be seen. 
The branch length estimates from either focal sequence (H1 and H3) are longer by the ExpCM than by the \texttt{phylobayes} mutation-selection model. 
This indicates that the DMS preferences are still more accurate for the local branches around the focal sequence than the global preferences estimated by \texttt{phylobayes}.

\skhcomment{What is the average difference in branch length between the two models?}

\begin{figure}[H]
\centerline{\includegraphics[width=0.5\textwidth]{figures/phylobayes.pdf}}
\caption{\label{fig:phylobayes}
\textbf{Comparison of ExpCM and phylobayes}
\skhcomment{y=x line too faint?}
}
\end{figure}

>>>>

The previous sections show that the ExpCM's, with their site-specific stationary states, estimate longer branches than the GY94. 
However, the ExpCM stationary state has a strong local effect as the amino-acid preferences from deep mutational scanning experiments are most accurate near their focal sequence. 
There are other models with site-specific stationary states which infer all of the parameters from the phylogenetic data rather than from independent experiments. 
For example, the CAT model, an amino-acid rather than codon model, describes the stationary state at each site as a mixture model of distinct, finite amino-acid profiles and the mutation-selection model implemented in \texttt{phylobayes} is almost identical to the ExpCM except the amino-acid preferences are inferred in a Bayesian framework rather than defined by deep mutational scanning experiments. 
We compared the ExpCM with the \texttt{phylobayes} mutation-selection model to examine the effect of site-specific stationary states on branch length estimation with and without experimental preferences.

The results in \ref{fig:phylobayes} show that the ExpCM+$\Gamma\omega$(H1+H3 avg prefs) and the \texttt{phylobayes} mutation-selection model, which includes $\Gamma\omega$, estimated similar branch lengths on the HA tree in \ref{fig:empirical_trees}. 
While the two models estimated the same branch length on average, the tension between local and global accuracy is still apparent in the results. 
All of the long branches from either the H1 or H3 focal sequence were estimated to be longer by the ExpCM+$\Gamma\omega$ while all other branches were estimated to be longer by the \texttt{phylobayes} mutation-selection model. 
This enrichment of branches from the focal sequence in the branches estimated to be longer by the ExpCM+$\Gamma\omega$(H1+H3 prefs) indicates that the ``global" stationary state inferred by \texttt{phylobayes} is not as accurate as the deep mutational scanning preferences for sequences close to the focal sequence. 
The accuracy of the deep mutational scanning preferences is sacrificed for more a universal stationary state. 
While the parameters inferred by the \texttt{phylobayes} model are a good compromise, they do not avoid the tension between the importance of a site-specific stationary state for long branch estimation and shifting preferences.

\begin{figure}[H]
\centerline{\includegraphics[width=0.5\textwidth]{figures/phylobayes.pdf}}
\caption{\label{fig:phylobayes}
\textbf{Comparison of ExpCM and \texttt{phylobayes}}
We estimated the branch lengths of the HA tree in \ref{fig:empirical_trees} using the mutation-selection model implemented in \texttt{phylobayes}. 
In comparison to ExpCM's, the \texttt{phylobayes} model infers the site-specific amino-acid preferences from the phylogenetic data rather than defining them \textit{a priori} from deep mutational scanning experiments. 
}
\end{figure}
